Hyperplane Elimination for Quickly Enumerating Local Optima

### 2015-10-27 Hyperplane Enumeration/main.tex
Minor edits from Bill

The first major result in Gray-Box optimization was the proof that the list
of fitness-improving moves from a solution can be updated after a bit flip in \BigO{1}
time~\cite{whitley:2012:constant}.
The fitness effect of flipping $x_j$ only changes after flipping $x_i$ if
there is a non-linear relationship between $x_i$ and $x_j$.
In an Mk Landscape, variables
can only have a non-linear relationship if they appear together in at least one $s_i$.
Therefore by definition the total number of non-linear
relationships in an Mk Landscape is linear with $N$, meaning on average the amortized
number of non-linear relationships per variable is \BigO{1}.

@@

This reduction in growth complexity translates into a substantial increase in the size of problems
which can be enumerated. For instance, using the previous best enumeration methods~\cite{ochoa:2015:crossovernetworks},
the largest Adjacent NKq problem size that can be enumerated in 4 hours is 34 bits. Given the same amount of time,
Hyper-Reorder can enumerate problems with 77 bits. Using the regression model, we estimate solving
a 77 bit problem using Gray-Box would require 900~million years. Similarly, Hyper-Reorder's largest
Random NKq size of 69 bits would require Gray-Box 3~million years. We believe these larger problems
are more likely to share characteristics with those where search algorithms are actually applied.

>>>>

The first major result in Gray-Box optimization was the proof that the list
of fitness-improving moves from a solution can be updated after a bit flip in \BigO{1}
time~\cite{whitley:2012:constant}.
The fitness effect of flipping $x_j$ only changes after flipping $x_i$ if
there is a non-linear relationship between $x_i$ and $x_j$.
In an Mk Landscape, variables
can only have a non-linear relationship if they appear together in at least one $s_i$.
By definition the total number of non-linear
relationships in an Mk Landscape is linear with $N$, meaning on average the amortized
number of non-linear relationships per variable is \BigO{1}.

@@

This reduction in growth complexity translates into a substantial increase in the size of problems
which can be enumerated in reasonable time. For instance, using the previous best enumeration methods~\cite{ochoa:2015:crossovernetworks},
the largest Adjacent NKq problem size that can be enumerated in 4 hours is 34 bits. Given the same amount of time,
Hyper-Reorder can enumerate problems with 77 bits. Using the regression model, we estimate solving
a 77 bit problem using Gray-Box would require 900~million years. Similarly, Hyper-Reorder's largest
Random NKq size of 69 bits would require Gray-Box 3~million years. We believe these larger problems
are more likely to share characteristics with those where search algorithms are actually applied.
